The Array of Long-Baseline Antennas for Taking Radio Observations from the Sub-Antarctic

### 2020-07-04 albatros_paper.tex
HCC added new waterfall plots and text for 2-element results

\section{Two-element pathfinder}

@@

To assess the loss of sensitivity as the signal level increases, we
use \autoref{eqn:1bit_output} to derive the correlator efficiency
$\epsilon$, which is the ratio of measured signal-to-noise ratio (SNR)
to the ideal, infinite-precision SNR.  \autoref{fig:1bit_efficiency}
illustrates that the decrease in $\epsilon$ as a function of SNR is
relatively shallow.  When the SNR is near zero, $\epsilon$ approaches
$2/\pi \approx 0.64$, and if the signal level increases to two times
the  
noise level, $\epsilon$ decreases by only a factor of $\sim2$.  The
final \albatros\ array will have baselines that are long relative to
the observing wavelength, and the correlated component of the electric
field (which sets the relevant signal level) is likely a small
fraction of the total received power.  This qualitative argument can
by made by considering antennas in the $uv$ plane in the flat-sky
approximation.  Any individual antenna receives all power present in
the $uv$ plane (plus system noise), but the correlated contribution
comes only from the portion of the $uv$ plane that is encompassed by
the primary beam at the antenna's $uv$-space coordinate.  Assuming
that the sky can be described as a Gaussian random field, the maximum
correlated signal between two antennas is determined by the aperture
filling factor of that pair.  For the case of dipole antennas, the
filling factor is the square of the wavelength divided by the baseline
length.  Therefore, in the long-baseline limit, the correlated signal
is small in comparison to the total power received, and the correlator
efficiency is unlikely to decrease significantly from the ideal
$2/\pi$ value.

@@

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{Figures/Raw-ALBATROS-autospectra.PNG}
        \caption{} \label{Fig:waterfall_auto}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.48\textwidth}
      \centering
        \includegraphics[width=\linewidth]{Figures/First-fringes-of-ALBATROS-EGG.PNG}
        \caption{} \label{Fig:waterfall_phase}
    \end{subfigure}
    \caption{{\bf (a)} text goes here {\bf (b)} text goes
      here}\label{Fig:waterfalls}
\end{figure}

@@

\attention{\begin{itemize}
    \item{Improved waterfall plots from 2-element pathfinder}
    \item{Plot showing that the solar power works (sort of)}
    \item{Some sort of...data...plot for the autonomous station.  Can
      we say anything about baseband performance?}
  \end{itemize}
}

\autoref{Fig:waterfall_auto} and \autoref{Fig:waterfall_phase} shows
the initial results from the two interferometric array.
\autoref{Fig:waterfall_auto} shows the raw ALBATROS-EGG autospectra
which where the waterfall plot was taken from one polarization (pol0)
over an interval of 3 days. The Galaxy rising/setting is clearly
visible in the structure. There are also ripples in frequency because
of uncalibrated data, and the ripples arise from reflections in the
cables. There is a qualitative difference between daytime and
nighttime data and this shows that the contamination from shortwave
radio drops off significantly at night, when the ionosphere becomes
quieter.

\autoref{Fig:waterfall_phase} shows the first fringes that were
detected by the the ALBATROS-EGG. The fringes show recurrent structure
down to a frequency of as low as 10 MHz without data processing or
data cuts.

>>>>

\section{Two-element pathfinder}\label{s:2elem}

@@

To assess the loss of sensitivity as the signal level increases, we
use \autoref{eqn:1bit_output} to derive the correlator efficiency
$\epsilon$, which is the ratio of measured signal-to-noise ratio (SNR)
to the ideal, infinite-precision SNR.  \autoref{fig:1bit_efficiency}
illustrates that the decrease in $\epsilon$ as a function of SNR is
relatively shallow.  When the SNR is near zero, $\epsilon$ approaches
$2/\pi \approx 0.64$, and if the signal level increases to two times
the  
noise level, $\epsilon$ decreases by only a factor of $\sim2$.  The
final \albatros\ array will have baselines that are long relative to
the observing wavelength, and the correlated component of the electric
field (which sets the relevant signal level) is likely a small
fraction of the total received power.  This qualitative argument can
by made by considering antennas in the $uv$~plane in the flat-sky
approximation.  Any individual antenna receives all power present in
the $uv$~plane (plus system noise), but the correlated contribution
comes only from the portion of the $uv$~plane that is encompassed by
the primary beam at the antenna's $uv$-space coordinate.  Assuming
that the sky can be described as a Gaussian random field, the maximum
correlated signal between two antennas is determined by the aperture
filling factor of that pair.  For the case of dipole antennas, the
filling factor is the square of the wavelength divided by the baseline
length.  Therefore, in the long-baseline limit, the correlated signal
is small in comparison to the total power received, and the correlator
efficiency is unlikely to decrease significantly from the ideal
$2/\pi$ value.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.9\linewidth]{Figures/albatros2_waterfalls.jpg}
    \caption{Spectra from a co-aligned polarization pair from the
      two-element \albatros\ pathfinder, shown as a function of
      frequency and time.  The auto- and cross-spectra are shown in
      the top and bottom rows, respectively.  The spectrum magnitudes
      are in uncalibrated ADC units on a logarithmic scale.
      Repeatable structure from the Galaxy is visible over the
      three-day time scale in all plots.}
    \label{Fig:albatros2_waterfalls}
  \end{center}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.6\linewidth]{Figures/albatros2_lst_phase.jpg}
    \caption{Phases from the cross-correlation between two co-aligned
      polarizations in the two-element \albatros\ pathfinder.  About
      372 hours of data are shown here, binned in local sidereal
      time.}
    \label{Fig:albatros2_phase}
  \end{center}
\end{figure}

@@

The primary goal of the two-element pathfinder described in
\S\ref{s:2elem} was to qualitatively assess the level of RFI and
ionospheric contamination at low frequenices.
\autoref{Fig:albatros2_waterfalls} illustrates auto- and cross-spectra
from one of the co-aligned polarization pairs from the two antennas.
The data are shown over a three-day period (June 18--21, 2018), and no
calibration or cuts have been applied.  The rising and setting of the
Galaxy is visible in all of the spectrum magnitudes, and repeatable
fringes are visible in the phases of the cross-spectra.  The RFI lines
at $\lesssim20$~MHz arise primarily from shortwave radio transmission
reflecting off the ionosphere.  During the night, when the ionospheric
plasma cutoff frequency drops, the maximum frequency of the RFI
contamination falls to $\sim10$~MHz.

As a qualitative illustration of the sky signal repeatability,
\autoref{Fig:albatros2_phase} shows the phases of the cross-spectra
binned in local sidereal time.  About 372 hours of data are averaged
into this plot, and no RFI excision has been performed.  A more
detailed analysis of signal repeatability will be presented in a
future paper, but the high signal-to-noise fringe pattern, which is
even visible a little below $\sim10$~MHz, demonstrates the proof of
concept for building the expanded \albatros\ array.

\attention{Do we want a plot showing solar power data of any sort?}

% \autoref{Fig:waterfall_auto} and \autoref{Fig:waterfall_phase} shows
% the initial results from the two interferometric array.
% \autoref{Fig:waterfall_auto} shows the raw ALBATROS-EGG autospectra
% which where the waterfall plot was taken from one polarization (pol0)
% over an interval of 3 days. The Galaxy rising/setting is clearly
% visible in the structure. There are also ripples in frequency because
% of uncalibrated data, and the ripples arise from reflections in the
% cables. There is a qualitative difference between daytime and
% nighttime data and this shows that the contamination from shortwave
% radio drops off significantly at night, when the ionosphere becomes
% quieter.
% 
% \autoref{Fig:waterfall_phase} shows the first fringes that were
% detected by the the ALBATROS-EGG. The fringes show recurrent structure
% down to a frequency of as low as 10 MHz without data processing or
% data cuts.
